---
title: "R Exercises"
output: html_notebook
---
```{r}
# source: http://rstatistics.net/
setwd("/Users/laurita/Documents/GitHub/ds_moocs/r_programing/")
```

# R Tutorial
This worksheet is meant to refresh R concepts and put them to practice in a near future. The source of the tutorial is: http://rstatistics.net/r-tutorial-exercise-for-beginners/

## Environments
The outermost main container is called the global environment (globalenv()).
This is the default place will R will store all objects that you create. You can place as many objects as your computer memory will allow it to hold.

```{r}
env1 <- new.env()  # create a new environment
assign("a", 3, envir = env1)  # store a=3 inside env1
ls()  # returns objects in .GlobalEnv
ls(env1)  # returns objects in env1
get('a', envir=env1)  # retrieve value from env1
```

Once you know how to insert objects into your environment, 
here is how to erase them.
```{r}
rm(a)  # delete the object 'a'
rm(list = ls())  # caution: delete all objects in .GlobalEnv
rm(list=ls())  # remove all objects in work space
#gc()  # free system memory
```

## Classes & Data types

### Variable Types
(https://www.tutorialspoint.com/r/r_data_types.htm)

#### character
Strings

#### integer
Integers

#### numeric
Integers + Fractions

#### factor
Categorical variable where each level is a category

#### logical
Boolean (e.g TRUE/FALSE)

#### complex
Complex numbers (3 + 2i)

#### raw
Storage coded (e.g "Hello" is stored as 48656c6c6f)

```{r}
l <- TRUE
l; class(l)

a <- 3.14
a; print(class(a))

b <- 'hello world'
b ; print(class(b))

d <- 2L
d ; class(d)
  
m <- 2+5i
m ; class(m)

v <- charToRaw("laura")
v; class(v)
```

### Data Types
There are many types of R-objects. The frequently used ones are:

#### Vector
In R programming, the very basic data types are the R-objects called *vectors* which hold elements of (same) variable classes of mode 'mode'. Please note in R the number of classes is not confined to only the above six types, e.g we can use many atomic vectors and create an array whose class will become array.

##### Usage
vector(mode = "logical", length = 0)

as.vector(x, mode = "any")

is.vector(x, mode = "any")

```{r}
# Create a vector.
apple <- c('red','green',"yellow")
print(apple); print(class(apple))
```

#### Matrix
A matrix is a two-dimensional rectangular data set (n X m). It can be created using a vector input to the matrix function. All columns must uniformly contain only one variable type.
```{r}
# Create a matrix.
M = matrix( c('a','a','b','c','b','a'), nrow = 2, ncol = 3, byrow = TRUE)
print(M); print(class(M))
````

#### Array
While matrices are confined to two dimensions, arrays can be of any number of dimensions. The array function takes a dim attribute which creates the required number of dimension. In the below example we create an array with two elements which are 3x3 matrices each. A two-dimensional array is the same thing as a matrix. One-dimensional arrays often look like vectors, but may be handled differently by some functions: str does distinguish them in recent versions of R.

Array returns an array with the extents specified in dim and naming information in dimnames. The values in data are taken to be those in the array with the leftmost subscript moving fastest. If there are too few elements in data to fill the array, then the elements in data are recycled. If data has length zero, NA of an appropriate type is used for atomic vectors (0 for raw vectors) and NULL for lists.

Unlike matrix, array does not currently remove any attributes left by as.vector from a classed list data, so can return a list array with a class attribute.
##### Usage ?array
```{r}
array(data = [vector], dim = length(data), dimnames = NULL)
as.array(x, ...)
is.array(x)
```

##### Arguments
* data: a vector (including a list or expression vector) giving data to fill the array. Non-atomic classed objects are coerced by as.vector.

* dim:  the dim attribute for the array to be created, that is an integer vector of length one or more giving the maximal indices in each dimension. (dim = n , m, z)

```{r}
# Create a list.
array1 <- array(c('green','yellow'), dim = c(3,3,2))
print(array1); print(class(array1))
````

#### Data.frame
Data frames are tabular data objects (n X m). Unlike a matrix, in data frame each column can contain different modes of data. The first column can be numeric while the second column can be character and third column can be logical. It is a list of vectors of equal length.

Data Frames are created using the data.frame() function.
```{r}
# Create a data frame.
BMI <- 	data.frame(
   gender = c("Male", "Male","Female"), 
   height = c(152, 171.5, 165), 
   weight = c(81,93, 78),
   Age = c(42,38,26))
print(BMI); print(class(BMI))
````

#### List
A list is an R-object which can contain many different types of elements inside it like vectors, functions and even another list inside it. Can hold objects of different classes and lengths.

```{r}
# Create a list.
list1 <- list(c(2,5,3),21.3,sin)
print(list1); print(class(list1))
````

#### Factor
Factors are the r-objects which are created using a vector. It stores the vector along with the distinct values of the elements in the vector as labels. The labels are always character irrespective of whether it is numeric or character or Boolean etc. in the input vector. They are useful in statistical modeling. Factors are created using the factor() function. The nlevels functions gives the count of levels.
```{r}
# Create a list.
apple_colors <- c('green','green','yellow','red','red','red','green')
factor_apple <- factor(apple_colors)
print(factor_apple); print(class(factor_apple))
````

